#Data preparation

##Load and prepare survey data

First we define where the data is stored, and subsequently define the libraries to load

```{r, echo=TRUE}
if (Sys.info()["nodename"]=="L0151542") {path<-"c:/github/CoastalFish/fish_book/data";
                        env.path<-"C:/Users/aarts012/Dropbox/Seal_database/environmental"}
library(INLA); library(fields); 
library(mgcv)
library(lattice); library(latticeExtra); library(grid); library(gridExtra);
library(rgdal); library(rgeos); 
library(mapdata); library(maptools);
library(raster);library(geosphere);
library(spatialEco); library(rgeos); 
library(RColorBrewer);library(GISTools)

```

Next we load the data and standardize it in a way that it can be used by a more the standardized R INLA protocol.

```{r dataload, echo=TRUE}

dat<-read.csv(file.path(path,"fish/dfs/dfs4pmr_aco_ple.csv"))
names(dat)<-tolower(names(dat))
dat$date<-paste(dat$day,"/",dat$month,"/",dat$year,sep="")
names(dat)[names(dat) == 'latitude_s'] <- 'lat'
names(dat)[names(dat) == 'longitude_s'] <- 'lon'
names(dat)[names(dat) == 'snumtot'] <- 'count'
dat$date<-as.POSIXct(strptime(dat$date,format="%d/%m/%Y"))

dat<-dat[is.na(dat$lat)==FALSE & is.na(dat$lon)==FALSE,]

xy<-project(cbind(dat$lon,dat$lat), "+init=epsg:32631")
dat$x_utm<-xy[,1]
dat$y_utm<-xy[,2]
  
head(dat)

```


Some processing needs to be done. This includes removing rows with no data for the coordinates. The NA's for snumtot/count are actualy zeros. However, because the proportion of a specific ages ends up in the offset, we set it to >0.1. Because the corresponding count = 0, it doesn't matter what value this is, as long as it is >0. It is not used anyway. 

```{r data_process, echo=TRUE}
dat<-dat[is.na(dat$lat)==FALSE & is.na(dat$lon)==FALSE,]
dat$count[is.na(dat$count)]<-0
for (age in 1:10){
  dat[,paste("age",age,sep="")][is.na(dat[,paste("age",age,sep="")])]<-0.000001
 }

```
Next we calculate the offset. The column stotnum/count represent the actual number of individuals. For the age, a age-length key is used to convert the raw counts to numbers per age class. Since we would like to maintain the raw data, we specify an offset. 

```{r offset, echo=TRUE}
dat$age.fraction<-dat$age1/dat$count
dat$age.fraction[dat$age.fraction<0.000001]<-0.000001
dat$age.fraction[dat$age.fraction>1e5]<-1e5
dat$the.offset<-log(1/dat$age.fraction)
dat$the.offset<-0
dat$count<-round(dat$age1)
```

We remove data from Germany and Denmark, and for the moment only concentrate on the data from 2017 for plaice.

```{r datalimit, echo=TRUE}
dat<-dat[is.element(dat$pmr_area_code,c("WaddenkustDE","WaddenkustDK"))==FALSE,]

dat.s<-dat[dat$year==2017,]

```

We load some land data to be used for plotting.

```{r landread, echo=TRUE}
library(rgdal)
land<-rgdal::readOGR(dsn=file.path(env.path,"landsea/WGS84/"),layer="WGS84geoKust500")
projection(land)<-CRS("+proj=longlat +datum=WGS84")
land.utm<-land
land.utm <- spTransform(land,CRS("+init=epsg:32631"))
```   
    

And plot Plaice age 0

```{r, eval=T, echo=TRUE, dpi=600, fig.width=8, fig.height=8}
plot(dat.s$lon,dat.s$lat, 
     cex=(0.2+log(dat.s$age0+1)/3), 
     col=c('black'),
     xlab="Longitude", ylab="Lattitude", main="Counts per hauls Age 0")
plot(land,add=TRUE,border="grey75",col="grey90")
box()
```

And plot Plaice age 1

```{r, eval=T, echo=TRUE, dpi=600, fig.width=8, fig.height=8}
plot(dat.s$lon,dat.s$lat, 
     cex=(0.2 +log(dat.s$age1+1)/1), 
     col=c('red'),
     xlab="Longitude", ylab="Lattitude", main="Counts per hauls Age 1")
plot(land,add=TRUE,border="grey75",col="grey90")
box()
```

Clearly there are large difference between the distribution of the two age classes

### The count data

Example of a sub-paragraph




