#Data preparation

##Load and prepare survey data

There are two file types; 
- Trawl table, defining where, when and how trawling took place
- Catch data, defining which species (and size) were caught

The trawl list table is the one to be linked to environmental data, and can later be linked to the catch data.

First the necessary file location 'paths' are specified.  

```{r, echo=TRUE}
if (Sys.info()["nodename"]=="L0151542") {path<-"c:/github/CoastalFish/fish_book/data";
env.path<-"C:/Users/aarts012/Dropbox/Seal_database/environmental";
PMR.data.path<-"W:/IMARES/Data/PMR/PMR 2019/Perceel Vis/3. data/"}
```

Libraries are loaded

```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(INLA); library(fields); 
library(mgcv)
library(lattice); library(latticeExtra); library(grid); library(gridExtra);
library(rgdal); library(rgeos); 
library(mapdata); library(maptools);
library(raster);library(geosphere);
library(spatialEco); library(rgeos); 
library(RColorBrewer);library(GISTools)
#library(vmstools); 
library(GISTools) ;library(ncdf4); 
library(RANN);

```

Next we load the trawl data and standardize it in a way that it can be used by a more the standardized R INLA protocol.

```{r trawlload, echo=TRUE}
dat<-read.csv(file.path(PMR.data.path,"DFS/dfs4pmr_trllst.csv"))
names(dat)<-tolower(names(dat))
```

A column for the date and time is created based on separate columns

```{r time_process_trawl, echo=TRUE}
dat$time<-sprintf("%04d", dat$time)
dat$date_time<-paste(dat$year,"-", dat$month,"-",dat$day," ", dat$time, sep="")
dat$ddate<-as.POSIXct(strptime(dat$date_time,format="%Y-%m-%d %H%M",tz="UTC")) 
tail(dat[,c("date_time","ddate")])
dat$day_of_year<-as.numeric(difftime(dat$ddate,as.POSIXct(strptime(paste(dat$year,"-01-01 0000", sep=""),format="%Y-%m-%d %H%M",tz="UTC"),units="days")))
```

All necessary columns are given a standard name, and utm coordinates are calculated. All rows with no lat or lon coordinates are removed. 

```{r process_trawl, echo=TRUE}

# Define common projection
  common.projection<-"+proj=utm +zone=31 +ellps=WGS84 +datum=WGS84 +units=m +no_defs +towgs84=0,0,0"
  # Same as +init=epsg:32631

# Change names
  names(dat)[names(dat) == 'latitude_s'] <- 'lat'
  names(dat)[names(dat) == 'longitude_s'] <- 'lon'
  
# Only select columns with data  
  dat<-dat[is.na(dat$lat)==FALSE & is.na(dat$lon)==FALSE,]

# Get utm coordinates  
  xy<-project(cbind(dat$lon,dat$lat), common.projection)
  dat$x_utm<-xy[,1]
  dat$y_utm<-xy[,2]
 
# Show data   
  str(dat)
``` 



