# Link trawl data (with environmental covariates) with catch data

### Load trawl data


```{r, echo=TRUE}
if (Sys.info()["nodename"]=="L0151542") {path<-"c:/github/CoastalFish/fish_book/data";
env.path<-"C:/Users/aarts012/Dropbox/Seal_database/environmental";
PMR.data.path<-"W:/IMARES/Data/PMR/PMR 2019/Perceel Vis/3. data/"}
```


```{r trawl_data, echo=TRUE, eval=T}

trawl<-read.csv(file.path(PMR.data.path,"DFS/dfs4pmr_trllst_ENV_2019-11-01.csv"))  
trawl$sampleID<-trawl$id

```

### Load catch data

```{r load_catch_data, echo=TRUE, eval=T}

catch<-read.csv(file.path(PMR.data.path,"DFS/dfs4pmr_catch.csv"))   

```

Define the species

```{r define species, echo=TRUE, eval=T}
the.species<-"Pleuronectes platessa"
#the.species<-"Platichthys flesus"
the.age<-0
```

Select only the valid trawls

```{r valid_trawls, echo=TRUE, eval=T}
table(trawl$valid, useNA="always")
n1<-nrow(trawl)
trawl <- trawl[trawl$valid == "y",]
n1-nrow(trawl)
```

Create sample_id for both the trawl data and the catch data.

```{r sampleID, echo=TRUE, eval=T}

# Sample id for catch data
  catch$sampleID <- paste(catch$year, catch$sample, sep="-")  ## different years share the same "sample" code
  length(unique(catch$sampleID))
  
```


Select only the trawl data where at least some individuals were caught. The idea is that those are valid trawls/hauls

```{r load_length_at_age, echo=TRUE, eval=T}
# Vector of unique sampleID  
trawl_list <- unique(trawl$sampleID)
catch_trawl_list <- unique(catch$sampleID)

n1<-nrow(trawl)
trawl<-trawl[is.element(trawl$sampleID,catch_trawl_list),]
n1-nrow(trawl)

n1<-nrow(catch)
catch<-catch[is.element(catch$sampleID,trawl_list),]
n1-nrow(catch)
```

```{r select species, echo=TRUE, eval=T}
catch<-catch[catch$SCIENTIFIC_NAME==the.species,]
```

Calculate age probabilities

```{r get_age_class, echo=TRUE, eval=T}

catch$age0.p<-1
catch$age1.p<-0
catch$lclass<-round(catch$LENGTH*100)

if (the.species=="Pleuronectes platessa") la<-read.csv(file.path(PMR.data.path,"DFS/dfs4pmr_overall_alk_ple.csv"))
if (the.species=="Solea solea") la<-read.csv(file.path(PMR.data.path,"DFS/dfs4pmr_overall_alk_sol.csv"))
if (the.species=="Platichthys flesus") la<-read.csv(file.path(PMR.data.path,"DFS/dfs4pmr_overall_alk_fle.csv"))
if (the.species=="Limanda limanda") la<-read.csv(file.path(PMR.data.path,"DFS/dfs4pmr_overall_alk_dab.csv"))

if (is.element(the.species,c("Pleuronectes platessa","Solea solea","Platichthys flesus","Limanda limanda"))){
               
la[is.na(la)]<-0
age.cols<-na.omit(match(paste("a",0:20,sep=""),names(la)))
la[,age.cols]<-la[,age.cols]/rowSums(la[,age.cols])
la$a1.plus<-rowSums(la[,age.cols[-1]])


catch$lclass.near<-la$class[apply(data.frame(catch[,"lclass"]),1,function(x){which.min(abs(x-la$class))[1]})]

catch$age0.p<-la$a0[match(catch$lclass.near,la$class)]
catch$age1.p<-la$a1.plus[match(catch$lclass.near,la$class)]
}


```

```{r numbers_per_haul, echo=TRUE, eval=T}

# Calculate the raw counts per sample ID
  tapply.n<-tapply(catch$QUANTITY,catch$sampleID,sum)
  
# Calculate the corrected (sub sampling factor) number per age
  if (the.age==0) tapply.na<-tapply(catch$number*catch$age0.p,catch$sampleID,sum) # 
  if (the.age==1) tapply.na<-tapply(catch$number*catch$age1.p,catch$sampleID,sum) # 

# Calculate the offset, which is the ratio between raw counts and corrected counts
  tapply.offset<-tapply.n/tapply.na

trawl$counts<-as.numeric(tapply.n[match(trawl$sampleID,names(tapply.n))])
trawl$the.offset<-as.numeric(tapply.offset[match(trawl$sampleID,names(tapply.offset))])
trawl$NatA<-as.numeric(tapply.na[match(trawl$sampleID,names(tapply.na))])

trawl$the.offset[trawl$the.offset=="Inf" | is.na(trawl$the.offset)]<-1
trawl$the.offset<-log(trawl$the.offset)
trawl$counts[is.na(trawl$counts)]<-0


```

Finally we remove data from Denmark and Germany

```{r, eval=T, echo=TRUE}
trawl<-trawl[trawl$lon<6,]
#trawl<-trawl[is.element(trawl$pmr_area_code,c("WaddenkustDE","WaddenkustDK"))==FALSE,]
```

Plot the counts

```{r, eval=T, echo=TRUE, dpi=600, fig.width=8, fig.height=8}
land<-rgdal::readOGR(dsn=file.path(PMR.data.path,"Environmental_variables/shapefiles/WGS84/"),layer="WGS84geoKust500")
projection(land)<-CRS("+proj=longlat +datum=WGS84")
trawl.s<-trawl[trawl$year==2018,]
plot(trawl.s$lon,trawl.s$lat, 
     cex=(0.2+log(trawl.s$NatA+1)/3), 
     col=c('black'),
     xlab="Longitude", ylab="Lattitude", main=paste(the.species, "  N/haul Age ",the.age,sep=""))
plot(land,add=TRUE,border="grey75",col="grey90")
box()
```

Finally we rename trawl to dat

```{r, eval=T, echo=TRUE}
dat<-trawl
```

### recategorize into eco groups
spe_group$SCIENTIFIC_NAME <- as.character(spe_group$SCIENTIFIC_NAME)
dat$SCIENTIFIC_NAME       <- as.character(dat$SCIENTIFIC_NAME)
table(spe_group$ecotype, useNA="always")
spe_group <- spe_group[spe_group$ecotype!="",]
dat       <- merge(dat,spe_group[,c("SCIENTIFIC_NAME", "ecotype")], by="SCIENTIFIC_NAME", all.x=T)
table(dat$ecotype, useNA="always")
table(dat$SCIENTIFIC_NAME[is.na(dat$ecotype)])
dat      <- dat[!is.na(dat$ecotype),]    ##delete spacies without categorization, e.g. freshwater species






